\documentclass[11pt,a4paper]{ctexart}
% ============================================================================
% 硕士选题报告 — 正经 LaTeX 模板(ctexart + biblatex/gb7714-2015 上标引用 + TikZ)
% 编译:xelatex -> biber -> xelatex -> xelatex
% ============================================================================
\usepackage[margin=2.4cm]{geometry}
\usepackage{xcolor}
\definecolor{thupurple}{RGB}{130,49,142}
\definecolor{thublue}{RGB}{0,78,152}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{float}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,fit,backgrounds}
\usepackage[backend=biber,style=gb7714-2015,sorting=none]{biblatex}
\addbibresource{../refs/refs.bib}
\usepackage[colorlinks,linkcolor=thupurple,citecolor=thupurple,urlcolor=thublue]{hyperref}

\setlist{nosep}
\linespread{1.18}
\newcommand{\zihaoCustom}{}
% 章节标题用黑体
\ctexset{section/format=\Large\bfseries\heiti, subsection/format=\large\bfseries\heiti}

\begin{document}
\sloppy
% ===================== 封面 =====================
\begin{center}
  {\zihao{2}\bfseries\heiti 基于大模型编码智能体的自进化 eBPF 网络入侵防御系统}\\[6pt]
  {\zihao{3}\heiti 硕士学位论文选题报告}\\[10pt]
  {\zihao{-4}\itshape A Self-Evolving eBPF Network Intrusion Prevention System\\ Driven by Large Language Model Coding Agents}\\[14pt]
\end{center}

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{0.92\textwidth}{>{\bfseries\heiti}l X >{\bfseries\heiti}l X}
\toprule
研究生 & 李兆曦 & 学科(专业) & 计算机科学与技术(0812) \\
研究方向 & 网络空间安全 / 网络协议安全 & 指导教师 & 徐恪 教授 \\
论文类型 & 应用研究 / 应用基础研究 & 课题来源 & 国家级科研项目 / 自选 \\
\multicolumn{2}{l}{} & 选题报告日期 & 2026 年 6 月 \\
\bottomrule
\end{tabularx}
\end{center}
\vspace{8pt}

% ===================== 摘要 =====================
\begin{center}\textbf{\heiti\large 摘\quad 要}\end{center}
\noindent 随着 TCP/IP 协议栈跨层交互歧义持续催生链外(off-path)协议攻击,而现有防御零散、滞后且高度依赖人工编写内核代码,本课题提出并构建一个由\textbf{大模型编码智能体}自主维护、运行于内核的\textbf{自进化 eBPF 网络入侵防御系统}。受“超越梯度的学习(Learning Beyond Gradients)”启发,本文提出\textbf{防御演化(Defensive Evolution)}范式:将防御能力沉淀为经内核验证器(verifier)证明安全的 eBPF 代码,而非神经网络权重;由编码智能体以新颖性与覆盖度引导的主动探索探查攻击面、进化式合成防御程序,并经“编译—验证—沙箱”可信闭环热加载到 XDP 数据面,从而持续、可审计、形式化可验证地进化防御能力。本文将作者在 off-path 协议攻击上积累的攻击者视角反向用于防御,构建攻防协同进化评测,量化系统对已知与未知协议攻击的防御效果、误报率与性能开销。

\smallskip
\noindent\textbf{\heiti 关键词:}网络协议安全;eBPF;大模型智能体;防御演化;自进化系统;入侵防御

\smallskip
\begin{center}\textbf{\large Abstract}\end{center}
\noindent Ambiguities in cross-layer TCP/IP interactions keep spawning off-path protocol attacks, yet existing defenses are fragmented, reactive, and rely on hand-written kernel code. This proposal designs a self-evolving, in-kernel \emph{eBPF network intrusion prevention system} autonomously maintained by Large Language Model coding agents. Inspired by ``learning beyond gradients'', we propose the \emph{Defensive Evolution} paradigm: defensive capability is accumulated as eBPF code proven safe by the kernel verifier rather than as neural-network weights. A coding agent explores the attack surface under novelty- and coverage-guided exploration, evolutionarily synthesizes defenses, and hot-loads them through a compile--verify--sandbox trusted loop, evolving the defense library continuously and auditably. Turning the author's offensive off-path expertise into defense, we build an attack--defense co-evolution evaluation to quantify protection against known and unknown attacks, false positives, and runtime overhead.

\smallskip
\noindent\textbf{Keywords:} network protocol security; eBPF; LLM agents; Defensive Evolution; self-evolving systems; intrusion prevention.

\vspace{6pt}\hrule\vspace{4pt}
{\small\tableofcontents}
\clearpage

% ===================== 正文 =====================
\input{body.tex}

% ===================== 参考文献 =====================
\printbibliography[title={参考文献}]
\addcontentsline{toc}{section}{参考文献}

\end{document}
